### 2011-07-20 Paper/Paper.tex
Updated paper, generated performances

@@

@@

Given a hypothetical perfect synthesizer, performing music with computers is a trivial task. However \textit{expressively} performing music is not and much research has focussed on this issue. Computers are widely used in music production. Since support for automatic expression is mostly absent or very poor in music editing software, some genres of music have evolved to sound acceptable even without expression. If automatic expression better in this software, computer music production could greatly widen its field of application.

@@

The goal is to find a `safe' method to segment music, that is, a method that does not generate segmentations that clash too much with human musical intuition. For this purpose the parse trees of delta onset (inter-onset interval, or simply IOI) and delta pitch (pitch interval) seem most suited: sudden jumps in pitch or onset often correspond to phrase transitions. In figure \ref{pitchonsettrees} a pitch and onset parse tree of the first few bars of etude 23 of Chopin are displayed.

\begin{figure}

\end{figure}
NOTE(we now use only onset on the mozart corpus, explain adaptive segmentation)

In the tree in figure \ref{pitchonsettrees} we can see that some nodes contain only subnodes, some nodes contain only notes and some nodes contain notes and subnodes. To translate the tree into a segmentation we do the following for for every node in the tree starting with the root node:
\begin{algorithm}
\label{alg:segmentation}
\begin{enumerate}
\item If the node is a note, at it as a singleton constituent to the segmentation
\item Expand the node.
\item If the node contains only subnodes, or less than $x$ notes, start from step 1 for every subnode and -note.
\item If the node contains at least $x$ notes, add the notes recursively contained by the current note to the segmentation.
\end{enumerate}
\end{algorithm}

The value of $x$ represents the tolerance of singleton constituents. We need this parameter because sometimes very long notes cause top level splits in the onset tree. We have found that setting it to the number of notes recursively contained (yielded) by the current node divided by 16 works quite well for works by Mozart.

@@

\begin{figure}
\includegraphics[scale=3]{img/melodysegments}
\caption{First four segments of piano sonata KV331 III by Mozart (grace notes are not shown)}
\end{figure}

NOTE (SECOND ORDER PITCH INTERVAL AND "SMOOTHED" IOI TREES FIGURES plus DISCUSSION)

@@

\subsection{Discretization}
To calculate probabilities, we must discretize the feature and parameter vectors.

@@

\begin{itemize}
\item Extract melody
\item Extract scorefeatures from melody
\item lookup notes in deviation, ignore missing notes
\item Extract expressive and non-expressive melody and extract expression features
\item Perform notes other than melody notes the same as the nearest melody note
\end{itemize}

@@

@@

\section{Future Work}
\begin{itemize}
\item Combine deltatrees
\item Use loudness and tempo direction instead of means (requires more data)
\item Generalize approach to incorporate hierarchical structure, let structure extend to note level, so note level expression and structurelevel expression become integrated
\end{itemize}
\subsection{Repetition}
The system would certainly benefit from a notion of repetition and similarity. Repetition is a very good indicator of constituent breaks. Repetition and similarity could also be used to improve expressiveness of performances. It is probably telling when a phrase is repeated three times and then slightly altered the fourth time. Although finding similarity and musically significant repetition is a subject of its own the delta framework could help to define repetition arbitrarily of transposition or rhythm. A list of pitch deltas can for example be used to detect repetition independend of transposition and a list of duration deltas can be used to detect repetition of rhythm independend of the notes used.

>>>>

%\usepackage{fullpage}

@@

\usepackage{subfigure}

@@

Given a hypothetical perfect synthesizer, performing music with computers is a trivial task. However \textit{expressively} performing music is not and much research has focussed on this issue. Computers are widely used in music production. Since support for automatic expression is mostly absent or very poor in music editing software, some genres of music have evolved to sound acceptable even without expression. If good working automatic expression was available in software, computer music production could greatly widen its field of application.

A performance rendering system is a computer program that automatically creates an expressive, humanlike performance of a musical score. Early attempts like Director Musices REF have tried to create use sets of rules to create performances. These systems
Expressively performing a musical score is known as performance rendering. A performance rendering system aims to create automate A yearly competition is held among performance rendering systems

@@

The goal is to find a `safe' method to segment music, that is, a method that does not generate segmentations that clash too much with human musical intuition. For this purpose the parse trees of delta onset (inter-onset interval, or simply IOI) and delta pitch (pitch interval) seem most suited: sudden jumps in pitch or onset often correspond to phrase transitions. These two trees are displayed in figure \ref{fig:pitchonsettrees}.

\begin{figure}
\centering
\subfigure[Delta onset tree]{
\includegraphics[scale=0.4]{img/onset_sym}
\label{fig:donset}
}
\subfigure[Second order delta pitch tree]{
\includegraphics[scale=0.3]{img/second_order_pitch_sym}
\label{fig:dpitch}
}
\caption{Parse trees of the first eight bars of piano sonata KV331 III. (Turkisch March)by Mozart }
\label{fig:pitchonsettrees}
\end{figure}

In the trees in figure \ref{fig:pitchonsettrees} we can see that some nodes contain only subnodes, some nodes contain only notes and some nodes contain notes and subnodes. To translate the tree into a segmentation we use algorithm \ref{alg:segmentation} starting with the root node.
\begin{algorithm}
\caption{Segmentation}
\label{alg:segmentation}
\begin{enumerate}
\item If the node is a note, at it as a singleton constituent to the segmentation
\item Expand the node.
\item If the node contains only subnodes, or less than $x$ notes, start from step 1 for every subnode and -note.
\item If the node contains at least $x$ notes, add the notes recursively contained by the current note to the segmentation.
\end{enumerate}
\end{algorithm}

The value of $x$ represents the tolerance of singleton constituents. We need this parameter because sometimes very long notes cause top level splits in the onset tree. We have found that setting it to the number of notes recursively contained (yielded) by the current node divided by 16 works quite well for works by Mozart. Our segmentation of the first eight bars of Mozart's Turkish March using a delta onset tree can be found in figure \ref{fig:mozartsegments}.\footnote{Note that if you use the delta tree from figure  \ref{fig:donset} you will not get this segmentation. This segmentation was based on a onset delta tree over the entire piece instead of just the first eight bars}

@@

\begin{figure}
\includegraphics[scale=3]{img/melodysegments}
\caption{First four segments of piano sonata KV331 III by Mozart (grace notes are not shown)}
\label{fig:mozartsegments}
\end{figure}

@@

\subsection{Discretization}

We chose to make the model discrete. We discretize expression and scorefeatures in a different way.

@@

%\begin{itemize}
%\item Extract melody
%\item Extract scorefeatures from melody
%\item lookup notes in deviation, ignore missing notes
%\item Extract expressive and non-expressive melody and extract expression features
%\item Perform notes other than melody notes the same as the nearest melody note
%\end{itemize}

@@

The segmentation was based only delta onset trees since this worked best for Mozart.

Since transition probabilities are not smoothed only transitions that actually occur in the corpus have a nonzero probability

@@

\section{Future Work}
\begin{itemize}
\item Combine deltatrees
\item Use loudness and tempo direction instead of means (requires more data)
\item Generalize approach to incorporate hierarchical structure, let structure extend to note level, so note level expression and structurelevel expression become integrated
\item Backoff smoothing
\item Staccato statistics
\end{itemize}
\subsection{Repetition}
The system would certainly benefit from a notion of repetition and similarity. Repetition is a very good indicator of constituent breaks. Repetition and similarity could also be used to improve expressiveness of performances. It is probably telling when a phrase is repeated three times and then slightly altered the fourth time. Although finding similarity and musically significant repetition is a subject of its own the delta framework could help to define repetition arbitrarily of transposition or rhythm. A list of pitch deltas can for example be used to detect repetition independend of transposition and a list of duration deltas can be used to detect repetition of rhythm independend of the notes used.
